Hypercoercions and a Framework for \\ Equivalence of Cast Calculi

### 2020-01-03 paper.tex
rewrote weak-bisim-between-S(C)-and-D and Correctness-of-S(C)

@@

@@

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values $v$ and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machines involve two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an evaluating
$\sOOinspect{e}{\mathcal{E}}{\kappa}$ or continuing $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observable ($o$). Observables include all value 
constructors and blame. The function converting values to 
observables ($observe(v) = o$) is defined in the obvious way.

@@

\begin{figure}
  Machine state and other runtime data structures
  \[
  \begin{array}{lrcl}
  \stxrule{\mathcal{E}}{Environments}{
  	\text{a partial function}\;\{\langle x , v \rangle , \dots \}
%  	\text{partial functions that map  variables to values}
  }
  \stxrule{v}{Values}{
    \vOOtt{} \mid
    \vOOfun{\mathcal{E}}{x}{e} \mid
    \vOOcons{v}{v} \mid
    \vOOinl{v} \mid
    \vOOinr{v} \mid   
    \vOOcast{v}{c}
  }
  \stxrule{c}{Casts}{
    \cOOcast{T}{l}{T}
  }
  \stxrule{I}{Injectable types (\lazyD)}{
    P
  }
  \stxrule{I}{Injectable types (\lazyUD)}{
    \POOunit \mid
    \POOfun{\star}{\star} \mid
    \star \times \star \mid
    \star + \star
  }
  \stxrule{o}{Observables}{
    \oOOinj \mid
    \oOOsole \mid
    \oOOfun \mid
    \oOOcons \mid
    \oOOinl \mid
    \oOOinr \mid
    \oOOblame{l}
  }
  \stxrule{r}{Cast results}{
    \rOOsucc{v} \mid
    \rOOfail{l}
  }
  \stxrule{s}{States}{
    \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
    \sOOreturn{v}{\kappa} \mid{}
    \sOOhalt{o}
  }
  \stxrule{\kappa}{Continuations}{
    \kOOmt \mid
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
    \kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa}
  }
  \stxrulecont{ 
    \kOOcdr{\kappa}\mid
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa}\mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa}
  }
  \stxrulecont{
    \kOOcaseIII{v}{v}{\kappa} \mid
    \kOOcast{c}{\kappa}
  }
  \end{array}
  \]

@@

  \redrule{
    \highlight{\sOOreturn{v}{
        \kOOcast{c}{\kappa}
      }}
  }{
\highlight{          
\begin{cases}
  \sOOreturn{v'}{\kappa} & \sidecond{applyCast(v,c) = \rOOsucc{v'}}
  \\
  \sOOhalt{\oOOblame{l}} & \sidecond{applyCast(v,c) = \rOOfail{l}}
\end{cases}}
  }{}
\redrule{
\sOOreturn{v}{\kOOmt}}{
\sOOhalt{observe(v)}}{}
  \end{array}
  \]

@@

@@

@@

%% Three observables on this definition leads to hypercoercion: 
%% \begin{enumerate}
%%  \item The length of normal coercion is at most three.
%%  \item Projections are always at the beginning when present.
%%  \item Injections and failures are always at the end when present.
%% \end{enumerate}

@@

\begin{definition}[Lazy D Cast ADT]
  \label{def:surely-lazyd}
  A  Cast is a Lazy D Cast if:
  \begin{enumerate}
  \item If $v : T$ then $applyCast(v,id(T)) = \mathtt{succ} \; v $
  \item If $\judgeType{v}{T_1}$ and
    $ \judgeTypeFT{c_1}{T_1}{T_2}$ and
    $ \judgeTypeFT{c_2}{T_2}{T_3}$\\
    then $applyCast(v,seq(c_1,c_2)) = 
    applyCast(v,c_1) \mbind{} \lambda v'.applyCast(v',c_2)$\\
    where 
    \[
    \begin{array}{rcl}
      \rOOsucc{v} \mbind{} f & = & f(v) \\
      \rOOfail{l} \mbind{} f & = & \rOOfail{l}
    \end{array}
    \]
  \item If $v : T_1$ and $T_1 \not\smile T_2$,
    then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
  \item If $v : \star$,
    then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
  \item If $v : P$,
    then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
    = applyCast(v,cast(P,l,P')) $
  \item If $v : P$,
    then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
  \item If $v : \POOunit$,
    then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
  \item If $(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
    then\\
    $ 
    applyCast(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}, 
    cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
    = 
    \rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\mathcal{E}}{x}{e}{seq(c_2,cast(T_2,l,T_4))})}$
  \item If $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
    \times 
    T_4))$\\
    $ = 
    \rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
    $ 
  \item If $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
  \item If $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\$
    applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
  \end{enumerate}
\end{definition}

@@

\begin{figure} 
  Build continuation \fbox{$wrap(k) = \kappa$}
  \[
  \begin{array}{rclc}
  \funrule{wrap(k)}{\kOOcast{id(T_1)}{k}}{
    \sidecond{k : T_1 \Longrightarrow T_2}}
  \end{array}
  \]

@@

  \hiredruleS{
    \sOOreturn{v}{\kOOcast{c}{k}}
  }{
  \begin{cases}
  \continue{k}{v'} & \sidecond{applyCast(v,c) = \rOOsucc{v'}} 
  \\
  \sOOhalt{\oOOblame{l}} & \sidecond{applyCast(v,c) = \rOOfail{l}}
  \end{cases}
  }{}
  \end{array}
  \]

@@

@@

\begin{definition}[Monoid]
  A Cast ADT is a monoid if 
  for all
  $c_1 : T_1 \Longrightarrow T_2$,
  $c_2 : T_2 \Longrightarrow T_3$, and
  $c_3 : T_3 \Longrightarrow T_4$,
  \begin{enumerate}
    \item $seq(id(T_1),c_1) = c_1$,
    \item $seq(c_1,id(T_2)) = c_1$, and
    \item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
  \end{enumerate}
\end{definition}

@@

\begin{figure}
%  $\text{\effCEK{C}} \approx \text{\ineffCEKD}$
% \[\dots\]
  \fbox{$s \approx s$}
  % Because the grammar for states of the two machines are not identical,
  % I think it is worthwhile to write out these rules explicitly. -Jeremy
  \begin{gather*}
    \inference{dom(\mathcal{E}_1) = dom(\mathcal{E}_2) \\
               \forall x {:} T \in dom(\mathcal{E}_1),\,
                   \mathcal{E}_1(x) \approx \mathcal{E}_2(x) : T\\
               \kappa_1 \approx \kappa_2 : T}
              {\langle e, \mathcal{E}_1, \kappa_1 \rangle \approx
              \langle e, \mathcal{E}_2, \kappa_2 \rangle}
    %% \quad
    %% \inference{v_1 \approx v_2 : T \quad \kappa_1 \approx \kappa_2 : T}
    %%           {\langle v_1, \kappa_1 \rangle  \approx
    %%           \langle v_2, \kappa_2 \rangle}
    \quad
    \inference{}
              {\sOOhalt{o} \approx \sOOhalt{o}} \\
  \end{gather*} 
  \fbox{$\kappa \approx \kappa : T$}
  \begin{gather*}
  \inference{
    k \approx \kappa : T
  }{
    \kOOcast{id(T)}{k} \approx \kappa
  }
  \quad
  \inference{
    \kOOcast{c}{k_1} \approx \kappa_2 : T_2
  }{
    \kOOcast{seq(cast(T_1,l,T_2),c)}{k_1} \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : I
  }{
    \hcvOOinj{I}{v_1} \approx \vOOcast{v_2}{\cOOcast{I}{l}{\TOOdyn}}
    : \TOOdyn
  }
  \\[1ex]
   %% This one is an obvious congruence rule. -Jeremy
   %% \inference{}{
   %%  \hcvOOtt \approx \vOOtt : \POOunit
   %% }
   %% \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\[1ex]
  \inference{
  	\Gamma , x : T_1 \vdash e : T_2
  	\\
  	\forall{} x \in dom(\Gamma).\;
  		\mathcal{E}_1(x) \approx \mathcal{E}_2(x) : \Gamma(x)
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\[1ex]
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2}}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\begin{array}{l}
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
\end{array}
\\[1ex]
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\[1ex]
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2}
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
  \begin{array}{l}
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  \end{array}
\\[1ex]
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\[1ex]
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}

The bisimulation relation for values is designed to make
Lemma~\ref{thm:lem-apply-cast-SD} true, that is, if $v_1 \approx v_2$,
then
  \[
  applyCast_{\text{\effCEK{C}}}(v_1,cast(T_1,l,T_2)) 
  \approx 
  applyCast_{\text{\ineffCEKD}}(v_2,\cOOcast{T_1}{l}{T_2})
  \]
So if two values of type $I$ are related, injecting them to $\TOOdyn$
gives related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a closure in \ineffCEKD, the closure in \effCEK{C} must
have identities as its casts.
%
If a closure on the left is related to a value, we can apply
equivalent casts on both sides to obtain related values.
%
The rules for products and sums are similar.
%
Lemma~\ref{thm:lem-apply-cast-SD} is used in our bisimulation proof to
handle the transitions that apply casts to values.

\begin{lemma}[$applyCast$ Preserves Bisimulation]
  \label{thm:lem-apply-cast-SD}
  Assume $C$ implements Lazy D Cast ADT,
  $v_1$ is a value in \effCEK{C},
  $v_2$ is a value in \ineffCEKD,
  and $v_1 \approx v_2 : T_1$,
  \[
  applyCast_{\text{\effCEK{C}}}(v_1,cast(T_1,l,T_2)) 
  \approx 
  applyCast_{\text{\ineffCEKD}}(v_2,\cOOcast{T_1}{l}{T_2})
  \]
\end{lemma}
\begin{proof}
  Case splitting whether $T_1$ is shallowly consistent with $T_2$.
  If yes, case splitting how they are shallowly consistent.
  Then apply properties (3) - (11) of Lazy D Cast when applicable.
\end{proof}

>>>>

\newcommand{\longdownmapsto}{\rotatebox[origin=c]{-90}{$\scriptstyle\longmapsto$}\mkern2mu}

@@

\newcommand{\judgeSreducePlus}[3]{#2 \longmapsto_{\mathcal{S}(#1)}^{+} #3}

@@

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values $v$ and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machines involve two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an evaluating
$\sOOinspect{e}{\mathcal{E}}{\kappa}$ or continuing $\sOOreturn{v}{\kappa}$ 
configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation ($o$). Observations include all value 
constructors and blame. The function converting values to 
observations ($observe(v) = o$) is defined in the obvious way.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

@@

\begin{figure}
  Machine state and other runtime data structures
  \[
  \begin{array}{lrcl}
  \stxrule{\mathcal{E}}{Environments}{
  	\text{a partial function}\;\{\langle x , v \rangle , \dots \}
%  	\text{partial functions that map  variables to values}
  }
  \stxrule{v}{Values}{
    \vOOtt{} \mid
    \vOOfun{\mathcal{E}}{x}{e} \mid
    \vOOcons{v}{v} \mid
    \vOOinl{v} \mid
    \vOOinr{v} \mid   
    \vOOcast{v}{c}
  }
  \stxrule{c}{Casts}{
    \cOOcast{T}{l}{T}
  }
  \stxrule{I}{Injectable types (\lazyD)}{
    P
  }
  \stxrule{I}{Injectable types (\lazyUD)}{
    \POOunit \mid
    \POOfun{\star}{\star} \mid
    \star \times \star \mid
    \star + \star
  }
  \stxrule{o}{Observations}{
    \oOOinj \mid
    \oOOsole \mid
    \oOOfun \mid
    \oOOcons \mid
    \oOOinl \mid
    \oOOinr \mid
    \oOOblame{l}
  }
  \stxrule{r}{Cast results}{
    \rOOsucc{v} \mid
    \rOOfail{l}
  }
  \stxrule{s}{States}{
    \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
    \sOOreturn{v}{\kappa} \mid{}
    \sOOhalt{o}
  }
  \stxrule{\kappa}{Continuations}{
    \kOOmt \mid
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
    \kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa}
  }
  \stxrulecont{ 
    \kOOcdr{\kappa}\mid
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa}\mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa}
  }
  \stxrulecont{
    \kOOcaseIII{v}{v}{\kappa} \mid
    \kOOcast{c}{\kappa}
  }
  \end{array}
  \]

@@

  \redrule{
    \highlight{\sOOreturn{v}{
        \kOOcast{c}{\kappa}
      }}
  }{
\highlight{          
\begin{cases}
  \sOOreturn{v'}{\kappa} & \sidecond{applyCast(v,c) = \rOOsucc{v'}}
  \\
  \sOOhalt{(\oOOblame{l})} & \sidecond{applyCast(v,c) = \rOOfail{l}}
\end{cases}}
  }{}
\redrule{
\sOOreturn{v}{\kOOmt}}{
\sOOhalt{observe(v)}}{}
  \end{array}
  \]

@@

\begin{proposition}[\ineffCEKD{} is deterministic]
	If $\judgeDreduce{s_1}{s_2}$ and $\judgeDreduce{s_1}{s_3}$ then $s_2 = s_3$.
\end{proposition}

@@

\begin{proposition}[\ineffCEKUD{} is deterministic]
	If $\judgeDreduce{s_1}{s_2}$ and $\judgeUDreduce{s_1}{s_3}$ then $s_2 = 
	s_3$.
\end{proposition}

@@

%% Three observations on this definition leads to hypercoercion: 
%% \begin{enumerate}
%%  \item The length of normal coercion is at most three.
%%  \item Projections are always at the beginning when present.
%%  \item Injections and failures are always at the end when present.
%% \end{enumerate}

@@

\begin{definition}[Lazy D Cast ADT]
  \label{def:Lazy-D-Cast-ADT}
  A  Cast is a Lazy D Cast if:
  \begin{enumerate}
  \item If $v : T$ then $applyCast(v,id(T)) = \mathtt{succ} \; v $
  \item If $\judgeType{v}{T_1}$ and
    $ \judgeTypeFT{c_1}{T_1}{T_2}$ and
    $ \judgeTypeFT{c_2}{T_2}{T_3}$\\
    then $applyCast(v,seq(c_1,c_2)) = 
    applyCast(v,c_1) \mbind{} \lambda v'.applyCast(v',c_2)$\\
    where 
    \[
    \begin{array}{rcl}
      \rOOsucc{v} \mbind{} f & = & f(v) \\
      \rOOfail{l} \mbind{} f & = & \rOOfail{l}
    \end{array}
    \]
  \item If $v : T_1$ and $T_1 \not\smile T_2$,
    then $applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
  \item If $v : \star$,
    then $applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
  \item If $v : P$,
    then $applyCast(\hcvOOinj{P}{v},cast(\star,l,P')) 
    = applyCast(v,cast(P,l,P')) $
  \item If $v : P$,
    then $applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
  \item If $v : \POOunit$,
    then $applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
  \item If $(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}) : \POOfun{T_1}{T_2}$,
    then\\
    $ 
    applyCast(\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2}, 
    cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
    = 
    \rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\mathcal{E}}{x}{e}{seq(c_2,cast(T_2,l,T_4))})}$
  \item If $(\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
    \times 
    T_4))$\\
    $ = 
    \rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
    $ 
  \item If $(\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\
    $ 
    applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
  \item If $(\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2}$,
    then \\$
    applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
    = \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
  \end{enumerate}
\end{definition}

@@

\begin{figure} 
  Build continuation \fbox{$wrap(k) = \kappa$}
  \[
  \begin{array}{rclc}
  \funrule{wrap(k)}{\kOOcast{id(T_1)}{k}}{
  	\text{where } k : T_1 \Longrightarrow T_2}
  \end{array}
  \]

@@

  \hiredruleS{
    \sOOreturn{v}{\kOOcast{c}{k}}
  }{
  \begin{cases}
  \continue{k}{v'} & \sidecond{applyCast(v,c) = \rOOsucc{v'}} 
  \\
  \sOOhalt{(\oOOblame{l})} & \sidecond{applyCast(v,c) = \rOOfail{l}}
  \end{cases}
  }{}
  \end{array}
  \]

@@

\begin{proposition}[\effCEK{C} is deterministic]
	\label{prp:SC-deterministic}
	If $\judgeSreduce{C}{s_1}{s_2}$ and $\judgeSreduce{C}{s_1}{s_3}$ 
	then $s_2 = s_3$.
\end{proposition}

@@

\begin{definition}[Monoid]
	\label{def:monoid}
  A Cast ADT is a monoid if 
  for all
  $c_1 : T_1 \Longrightarrow T_2$,
  $c_2 : T_2 \Longrightarrow T_3$, and
  $c_3 : T_3 \Longrightarrow T_4$,
  \begin{enumerate}
    \item $seq(id(T_1),c_1) = c_1$,
    \item $seq(c_1,id(T_2)) = c_1$, and
    \item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
  \end{enumerate}
\end{definition}

@@

\begin{figure}
%  $\text{\effCEK{C}} \approx \text{\ineffCEKD}$
% \[\dots\]

\fbox{$\Gamma \vdash \mathcal{E} \approx \mathcal{E}$}
\begin{gather*}
\inference{
}{
	\emptyset \vdash \emptyset \approx \emptyset
}
\quad
\inference{
	\Gamma \vdash \mathcal{E}_1 \approx \mathcal{E}_2 \\
	v_1 \approx v_2 : T
}{
	\Gamma , x : T \vdash 
		\mathcal{E}_1 [ x := v_1 ] \approx \mathcal{E}_2 [ x := v_2 ]
}
\end{gather*}

  \fbox{$s \approx s$}
  % Because the grammar for states of the two machines are not identical,
  % I think it is worthwhile to write out these rules explicitly. -Jeremy
  \begin{gather*}
    \inference{
    	\judgetype{\Gamma}{e}{T} \\
    	\Gamma \vdash \mathcal{E}_1 \approx \mathcal{E}_2 \\
   		\kappa_1 \approx \kappa_2 : T
   	}{
   		\langle e, \mathcal{E}_1, \kappa_1 \rangle 
   		\approx
		\langle e, \mathcal{E}_2, \kappa_2 \rangle
	}
%    \quad
%    \inference{v_1 \approx v_2 : T \quad \kappa_1 \approx \kappa_2 : T}
%               {\langle v_1, \kappa_1 \rangle  \approx
%               \langle v_2, \kappa_2 \rangle}
    \quad
    \inference{}
              {\sOOhalt{o} \approx \sOOhalt{o}} \\
  \end{gather*} 
  \fbox{$\kappa \approx \kappa : T$}
  \begin{gather*}
  \inference{
    k \approx \kappa : T
  }{
    \kOOcast{id(T)}{k} \approx \kappa
  }
  \quad
  \inference{
    \kOOcast{c}{k_1} \approx \kappa_2 : T_2
  }{
    \kOOcast{seq(cast(T_1,l,T_2),c)}{k_1} \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : I
  }{
    \hcvOOinj{I}{v_1} \approx \vOOcast{v_2}{\cOOcast{I}{l}{\TOOdyn}}
    : \TOOdyn
  }
  \\[1ex]
   %% This one is an obvious congruence rule. -Jeremy
   %% \inference{}{
   %%  \hcvOOtt \approx \vOOtt : \POOunit
   %% }
   %% \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\[1ex]
  \inference{
  	\Gamma , x : T_1 \vdash e : T_2
  	\\
  	\Gamma \vdash \mathcal{E}_1 \approx \mathcal{E}_2
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\[1ex]
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2}}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\begin{array}{l}
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
\end{array}
\\[1ex]
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\[1ex]
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2}
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
  \begin{array}{l}
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  \end{array}
\\[1ex]
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\[1ex]
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}

The bisimulation relation for values is designed to make
Lemma~\ref{lem:lem-apply-cast-preserve-bisim} true, that is, if $v_1 \approx 
v_2$,
then
  \[
  applyCast_{\text{\effCEK{C}}}(v_1,cast(T_1,l,T_2)) 
  \approx 
  applyCast_{\text{\ineffCEKD}}(v_2,\cOOcast{T_1}{l}{T_2})
  \]
So if two values of type $I$ are related, injecting them to $\TOOdyn$
gives related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a closure in \ineffCEKD, the closure in \effCEK{C} must
have identities as its casts.
%
If a closure on the left is related to a value, we can apply
equivalent casts on both sides to obtain related values.
%
The rules for products and sums are similar.
%
Lemma~\ref{lem:lem-apply-cast-preserve-bisim} is used in our bisimulation proof 
to
handle the transitions that apply casts to values.

\begin{lemma}[$applyCast$ Preserves Bisimulation]
  \label{lem:lem-apply-cast-preserve-bisim}
  Assume $C$ implements Lazy D Cast ADT,
  $v_1$ is a value in \effCEK{C},
  $v_2$ is a value in \ineffCEKD,
  and $v_1 \approx v_2 : T_1$,
  \[
  applyCast_{\text{\effCEK{C}}}(v_1,cast(T_1,l,T_2)) 
  \approx 
  applyCast_{\text{\ineffCEKD}}(v_2,\cOOcast{T_1}{l}{T_2})
  \]
\end{lemma}
\begin{proof}
  Case splitting whether $T_1$ is shallowly consistent with $T_2$.
  If yes, case splitting how they are shallowly consistent.
  Then apply properties (3) - (11) of Lazy D Cast when applicable.
\end{proof}
